---
title: "Understanding Fatigue For Grand Slam Matches"
subtitle: "Summer Research with Tennis Australia"
author: "Sherry Zhang"
date: "08/02/2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r echo = FALSE, result = FALSE, message=FALSE}
library(ggplot2)
library(tidyverse)
library(purrr)
library(broom)
library(gridExtra)




source("R/ExtractRecordall.R")
source("R/importance.R")
source("R/rest.R")
```


# A bit about myself...

- Bachelor of Commerce (2018)
- Econometrics 
- Honours degree in 2019 


---
# Fatigue in Tennis Players 

## Motivation

- psychological effect 
- crucial to win 

Thus, 

We want to model fatigue via an observable measure: Serve Speed 

???
this is notes? 

---

# Data
- Public data | 2014-2017 | Four grand slams | Male & Female

- Matches & Points

- Mainly experiment with 2017 ausopen data

- Alexander Zverev (Sasha)

--

## Variables
- Point indicators: match_num, SetNo, GameNo, PointNumber

- Speed_KMH

- ServeNo: 1st or 2nd serve 

- Dist, Time, RallyCount

---

## New Variables 
- brk1, brk2: [dummies] 
  - Whether the player takes a 90 second bi-game break or a 120 second set break
  
- impt, impt_high, impt_low: Point Importance 
  - given the current score in the game, how important the current point to the outcome of the set
  - based on Steph's work in `deuce` package

```{r echo = FALSE, message=FALSE, fig.height=3, dev='svg'}
Alex =extract_records_all("data/2017-ausopen-matches.csv",
                          "data/2017-ausopen-points.csv", "Alexander Zverev")

Alex <- Alex %>% filter(Speed_KMH != 0)

Alex["impt"] = point_impt(Alex)
Alex["rest"] = rest(Alex)


f <- data.frame(brk1 = ifelse(rest(Alex) == 1.5, 1,0), 
                brk2 = ifelse(rest(Alex) == 2, 1,0), 
                impt = Alex$impt, 
                t = Alex$PointNumber, 
                match_num = Alex$match_num,
                speed = Alex$Speed_KMH,
                GameNo = Alex$GameNo,
                ServeNo = Alex$ServeNumber)
f <- f %>% replace_na(list(impt = 0))

ggplot(f) + geom_histogram(aes(x = impt))

f <- f %>% filter(match_num == 1310) %>% filter(speed != 0 )%>% 
  mutate(impt_low = ifelse(impt<0.1, 1,0)) %>% 
  mutate(impt_high = ifelse(impt>=0.1, 1,0))

f <- f %>% mutate(t_n = (t-min(t))/(max(t)-min(t))*100) # treat point index to be between 0 and 1 to avoid computation problem in the exponential of concentration function 

f <- f %>% 
  mutate(firstserve_index = ifelse(ServeNo ==1, 1,0 ))


```
 
```{r eval = FALSE}
Alex <- Alex %>% mutate(impt_low = ifelse(impt<0.1, 1,0)) 
```
 



---
# Linear Model 

Separate models for first and second serve because

```{r echo = FALSE, fig.height=5.5,fig.width=10, message=FALSE, dev='svg'}
p1 <- ggplot(Alex, aes(x = as.factor(ServeNumber), fill = match_num)) + 
  geom_bar(position = "dodge") + 
  xlab("Serve Number") + 
  theme(legend.position = "none", text = element_text(size = 10)) + 
  ggtitle("Count of 1st and 2nd serve") 

p2 <- ggplot(Alex,aes(x=Speed_KMH,col = match_num)) +
  ylim(min(Alex$Speed_KMH),max(Alex$Speed_KMH))+
  geom_density() +
  facet_wrap(~ServeNumber, ncol = 1, scales = "free_x") + 
  scale_y_continuous("Frequency", breaks = scales::pretty_breaks(n = 5)) +
  scale_x_continuous("Speed (KPH)", breaks = scales::pretty_breaks(n = 10)) + 
  theme(legend.position = "right", text = element_text(size = 10)) + 
  ggtitle("Density of Serving Speed")

grid.arrange(p1,p2, nrow = 1)
```


---

```{r echo = FALSE, fig.height=7,fig.width=10, message=FALSE, dev='svg'}
ggplot(Alex,aes(y=Speed_KMH,x = PointNumber)) +
  geom_point() +
  geom_smooth() + 
  facet_grid(ServeNumber~match_num,  scales = "free_x") + 
  scale_y_continuous("Serve Speed", breaks = scales::pretty_breaks(n = 5)) +
  scale_x_continuous("Point Number", breaks = scales::pretty_breaks(n = 10)) + 
  theme(legend.position = "right", text = element_text(size = 10)) + 
  ggtitle("Serve Speed - Point Number")
```


---
```{r echo = FALSE}
fit_lm <- function(data) lm(Speed_KMH~ 
                              PointNumber+ impt + time + dist + rest + MatchNo + RallyCount,
                            data = data)


build_linear_model <- function(dt, fit){
  by_player_fit <- dt %>% 
    group_by(ServeNumber) %>% 
    nest() %>% 
    mutate(model = map(data, fit)) 
  return(by_player_fit)
}

fetch_coef <- function(dt){
  player_coef_fit <- dt %>%
    unnest(model %>% map(tidy)) %>%
    dplyr::select(ServeNumber, term,estimate) %>%
    spread(term, estimate)
}

# split by first and second serve 
Alex_split<- split(Alex,Alex$ServeNumber)
FirstServe <- Alex_split[[1]] 
SecondServe <- Alex_split[[2]]

# model for first serve for all four players
firstserve_lm <- build_linear_model(FirstServe, fit_lm) %>% 
  mutate(servenumber =as.factor(1)) 
firstserve_coef <- firstserve_lm %>% fetch_coef()
firstserve_aug_lm <- firstserve_lm %>% unnest(model %>% map(augment))
firstserve_fit_lm <- firstserve_lm %>% unnest(model%>% map(glance))


# model for first serve for all four players
secondserve_lm <- build_linear_model(SecondServe, fit_lm) %>%
  mutate(servenumber =as.factor(2))
secondserve_coef <- fetch_coef(secondserve_lm)
secondserve_aug_lm <- secondserve_lm %>% unnest(model %>% map(augment))
secondserve_fit_lm<- secondserve_lm %>% unnest(model%>% map(glance))

# augment first and second serve data together 
model_lm<- rbind(firstserve_aug_lm, secondserve_aug_lm) 


```


```{r eval = FALSE}
fit_lm <- function(data) lm(Speed_KMH~ 
                              PointNumber+ impt + time + dist + 
                              brk1 + brk2 + MatchNo + RallyCount, 
                            data = data)
```


```{r echo = FALSE, fig.height=5, fig.width = 10, dev='svg'}
# linear model 
ggplot(model_lm) +
  facet_wrap(~MatchNo, scale = "free_x") + 
  geom_point(aes(x = PointNumber, y = .fitted, col = as.factor(ServeNumber)), size = 0.8) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45)) +
  scale_y_continuous(limits = c(100,220)) + 
  ggtitle("Alex-lm")

```

We observe fatigue!

---

class: inverse, center, middle

# Dose Response Model

---


## Concentration: $conc$
Overall, we expect the serving speed of a player to follow an exponential decay.

$$d_0e^{t\phi_0}$$

```{r echo = FALSE, fig.height=3.5, fig.width=10, dev="svg"}
conc <- function(t, phi,d){
  conc <- d * exp(t * phi)
  return(conc)
}

# baseline spec
t <- seq(1,100, 1)
phi_0 <- -0.01
d_0 <- c(rep(3, 100))
conc_0 <- conc(t, phi_0, d_0)
conc_plot<- data.frame(t, conc_0)
ggplot(conc_plot) + 
  geom_line(aes(t, conc_0)) + 
  ylab("conc")

```

---
When there is a 90 second break, players are injected a dose, which also follows an exponential decay function. 

```{r echo = FALSE, fig.height=7, fig.width=10, dev="svg"}
phi_1 <- -1.5
d_brk1 <- c(rep(5,100))
conc_brk1 <- conc(t, phi_1, d_brk1)
brk1 <- c(c(0,0,0,0,1),rep(0, 95))
conc_brk1_s = c(rep(0, 5),conc_brk1)[1:length(brk1)]
conc_plot_1 <- cbind(conc_plot,conc_brk1_s)
ggplot(conc_plot_1) + geom_line(aes(x = t, y = conc_0 + conc_brk1_s)) + 
  ylab("conc")
```

---
## Roll out to all the 90 second break...

```{r echo = FALSE, fig.height=6, fig.width=10, dev="svg"}

# brk1&2 spec
brk1 <- rep(c(0,0,0,0,1),20)

# compute lagged expoential for breaks
brk1_index <- which(brk1 %in% 1)
conc_brk1_whole = data.frame(matrix(nrow = 100, ncol =length(brk1_index)))
for (i in 1: length(brk1_index)){
  conc_brk1_whole[,i] = c(rep(0, brk1_index[i]),conc_brk1)[1:length(conc_brk1)]
}
conc_brk1_sum <- rowSums(conc_brk1_whole)


# plot
conc_plot_1_whole <- data.frame(t, conc_0, conc_brk1_sum)
ggplot(conc_plot_1_whole) + geom_line(aes(x = t, y = conc_0 + conc_brk1_sum)) +
  ylab("conc")
```

---
## Roll out to 120 second break...

```{r echo = FALSE, fig.height=6, fig.width=10, dev="svg"}
# plot 
conc <- function(t, phi,d){
  conc <- d * exp(t * phi)
  df <- tibble(t= t, conc = conc)
  return(conc)
}

# baseline spec
t <- seq(1,100, 1)
phi_1 <- -0.01
phi_2 <- -1.5

d_3 <- c(rep(3, 100))
d_3.2 <- c(rep(3.2,100))
d_5 <- c(rep(5, 100))

conc_3 <- conc(t, phi_1, d_3)
conc_3.2 <- conc(t, phi_2, d_3.2)
conc_5 <- conc(t, phi_2, d_5)

conc_0<- data.frame(t, conc_3, conc_3.2, conc_5)


# brk1&2 spec
brk1 <- rep(c(0,0,0,0,1),20)
brk2 <- rep(c(rep(0, 19),1),5)
brk1 <- ifelse(brk2==1, 0, brk1)

# compute lagged expoential for breaks
brk1_index <- which(brk1 %in% 1)
conc_brk1 = data.frame(matrix(nrow = 100, ncol =length(brk1_index)))
for (i in 1: length(brk1_index)){
  conc_brk1[,i] = c(rep(0, brk1_index[i]),conc_3.2)[1:length(conc_3.2)]
}
conc_brk1_sum <- rowSums(conc_brk1)


brk2_index <- which(brk2 %in% 1)
conc_brk2 = data.frame(matrix(nrow = 100, ncol =length(brk2_index)))
for (i in 1: length(brk2_index)){
  conc_brk2[,i] = c(rep(0, brk2_index[i]),conc_5)[1:length(conc_5)]
}
conc_brk2_sum <- rowSums(conc_brk2)

# plot
conc_plot <- cbind(conc_0, conc_brk1_sum, conc_brk2_sum)
ggplot(conc_plot) + geom_line(aes(x = t, y = conc_3 + conc_brk1_sum + conc_brk2_sum)) + 
  ylab("conc")
```


---
# Model Specification

```{r eval = FALSE}
ll <- function(par, data){
  conc_1 = par[4]*(exp(f$t_n*par[1]))
  conc_2 = par[5]*(exp(f$t_n*par[2]))
  conc_3 = (par[5] + par[6])*(exp(f$t_n*par[3]))
  
  for (i in 1: length(brk1_index)){
    conc_brk1[,i] = c(rep(0,brk1_index[i]),conc_2)[1:nrow(f)]}
  conc_brk1_sum <- rowSums(conc_brk1)
  
  for (i in 1: length(brk2_index)){
    conc_brk2[,i] = c(rep(0,brk2_index[i]),conc_3)[1:nrow(f)]}
  conc_brk2_sum <- rowSums(conc_brk2)
  
  ss_pred = par[7] + conc_1 + conc_brk1_sum + conc_brk2_sum+ 
    par[8]*f$impt_high +par[9]*f$firstserve_index
}
```

- Use MLE to optimise

```{r echo = FALSE, result = FALSE}
# brk1&2 spec
brk1 <- f$brk1
brk2 <- f$brk2
brk1_index <- which(brk1 %in% 1)
conc_brk1 = data.frame(matrix(nrow = nrow(f), ncol =length(brk1_index)))
brk2_index <- which(brk2 %in% 1)
conc_brk2 = data.frame(matrix(nrow = nrow(f), ncol =length(brk2_index)))


ll <- function(par, data){
  conc_1 = par[4]*(exp(f$t_n*par[1]))
  conc_2 = par[5]*(exp(f$t_n*par[2]))
  conc_3 = (par[5] + par[6])*(exp(f$t_n*par[3]))
  
  for (i in 1: length(brk1_index)){
    conc_brk1[,i] = c(rep(0,brk1_index[i]),conc_2)[1:nrow(f)]}
  conc_brk1_sum <- rowSums(conc_brk1)
  
  for (i in 1: length(brk2_index)){
    conc_brk2[,i] = c(rep(0,brk2_index[i]),conc_3)[1:nrow(f)]}
  conc_brk2_sum <- rowSums(conc_brk2)
  
  ss_pred = par[7] + conc_1 + conc_brk1_sum + conc_brk2_sum+ 
    par[8]*f$impt_high +par[9]*f$firstserve_index
  
  u = f$speed - ss_pred
  f <- -sum(log(dnorm(u, par[10], par[11])))
  return(f)
}


fit_1 <- optim(c(-0.9,-5,-5,10,10,10, 
                 160, 10, 10, 5,20), 
               ll,method= "L-BFGS-B", data = f,
               lower = c(-0.09, -Inf, -Inf, 0,0,0, -Inf, -Inf, -Inf, -Inf,-Inf), 
               upper = c(-0.0000001,0,0,20,20,20,Inf, Inf, Inf, Inf, Inf))
# fit_1 <- optim(c(-0.9,-1.5,-1.5,5,5,5,180,10,10,10,30,10,20), 
#                ll,method= "L-BFGS-B", data = f,
#                lower = c(-0.09, -5, -5, 0,0,0, -Inf, -Inf,-Inf, -Inf, -Inf, -Inf,-Inf), 
#                upper = c(-0.0000001,0,0,20,20,20,Inf, Inf,Inf,Inf, Inf, Inf, Inf))

# fit_1$par

```

```{r  echo = FALSE, result = FALSE}
conc_1<- fit_1$par[4]*exp(f$t_n*fit_1$par[1])
conc_2<- fit_1$par[5]*exp(f$t_n*fit_1$par[2]) 
conc_3 <- (fit_1$par[5] + fit_1$par[6])*exp(f$t_n*fit_1$par[3]) 

conc_exp <- cbind(f, conc_1, conc_2,conc_3)


# compute lagged expoential for breaks
for (i in 1: length(brk1_index)){
  conc_brk1[,i] = c(rep(0, brk1_index[i]),conc_2)[1:length(conc_2)]
}
conc_brk1_sum <- rowSums(conc_brk1)


for (i in 1: length(brk2_index)){
  conc_brk2[,i] = c(rep(0, brk2_index[i]),conc_3)[1:length(conc_3)]
}
conc_brk2_sum <- rowSums(conc_brk2)


conc_plot<- cbind(conc_exp, conc_brk1_sum, conc_brk2_sum)
conc_plot <- conc_plot %>% 
  mutate(base = fit_1$par[7] + conc_1 +fit_1$par[9]*f$firstserve_index) %>% 
  mutate(base_brk1 = fit_1$par[7] + conc_1 +conc_brk1_sum + 
           fit_1$par[9]*f$firstserve_index) %>% 
  mutate(base_brk2 = fit_1$par[7] + conc_1 +conc_brk2_sum + 
           fit_1$par[9]*f$firstserve_index) %>% 
  mutate(ss_pred_no_impt = fit_1$par[7] + conc_1 +conc_brk1_sum + conc_brk2_sum + 
           fit_1$par[9]*f$firstserve_index) %>% 
  mutate(ss_pred_w_impt = fit_1$par[7] + conc_1 +conc_brk1_sum + conc_brk2_sum + 
           fit_1$par[8]*f$impt_high + 
           fit_1$par[9]*f$firstserve_index) 

```

---

class: inverse, center, middle

# Result

---

##  Baseline decay
```{r echo = FALSE, message=FALSE, fig.height=6, fig.width=10, dev="svg"}
ggplot(conc_plot, aes(x = t_n)) + 
  geom_line(data = subset(conc_plot, ss_pred_w_impt >170), aes(y = base)) + 
  geom_line(data = subset(conc_plot, ss_pred_w_impt <170), aes(y = base)) + 
  xlab("t")

```

---
## Adding 90 second break ...
```{r echo = FALSE, message=FALSE, fig.height=6, fig.width=10, dev="svg"}
ggplot(conc_plot, aes(x= t_n)) + 
  geom_line(data = subset(conc_plot, ss_pred_no_impt >170), aes(y = base_brk1)) +
  geom_line(data = subset(conc_plot, ss_pred_no_impt <170), aes(y = base_brk1)) +
  geom_vline(data = filter(conc_plot, brk1==1),
             aes(xintercept = t_n, alpha = 0.5), col = "red", linetype = "dashed") +
  geom_vline(data = filter(conc_plot, brk2 ==1), 
             aes(xintercept = t_n, alpha = 0.5), col = "blue") + 
  theme(legend.position = "none") + 
  xlab("t") + ylab("base_1")
```

---
## Adding 120 second break...
```{r echo = FALSE, message=FALSE, fig.height=6, fig.width=10, dev="svg"}
ggplot(conc_plot, aes(x= t_n)) + 
  geom_line(data = subset(conc_plot, ss_pred_w_impt >175), aes(y = base), linetype = "dashed") +
  geom_line(data = subset(conc_plot, ss_pred_w_impt <175), aes(y = base), linetype = "dashed") +
  geom_line(data = subset(conc_plot, ss_pred_w_impt >175), aes(y = ss_pred_no_impt)) +
  geom_line(data = subset(conc_plot, ss_pred_w_impt <175), aes(y = ss_pred_no_impt)) +
  geom_vline(data = filter(conc_plot, brk1==1),
             aes(xintercept = t_n, alpha = 0.5), col = "red", linetype = "dashed") +
  geom_vline(data = filter(conc_plot, brk2 ==1), 
             aes(xintercept = t_n, alpha = 0.5), col = "blue") + 
  theme(legend.position = "none") + 
  xlab("t") + ylab("conc")

```

---
## Adding point importance

```{r echo = FALSE, message=FALSE, fig.height=6, fig.width=10, dev="svg"}
ggplot(conc_plot, aes(x= t_n)) + 
  geom_line(data = subset(conc_plot, ss_pred_w_impt >175), aes(y = base), linetype = "dashed")+
  geom_line(data = subset(conc_plot, ss_pred_w_impt <175), aes(y = base), linetype = "dashed")+
  geom_line(data = subset(conc_plot, ss_pred_w_impt >175), aes(y = ss_pred_w_impt)) +
  geom_line(data = subset(conc_plot, ss_pred_w_impt <175), aes(y = ss_pred_w_impt)) +
  geom_vline(data = filter(conc_plot, brk1==1),
             aes(xintercept = t_n, alpha = 0.5), col = "red", linetype = "dashed") +
  geom_vline(data = filter(conc_plot, brk2 ==1), 
             aes(xintercept = t_n, alpha = 0.5), col = "blue") + 
  theme(legend.position = "none") + 
  xlab("t") + ylab("Speed_pred")

```

---
# Goodness of fit 

```{r echo = FALSE, message=FALSE, fig.height=4, fig.width=10, dev="svg"}
p1 <- ggplot(conc_plot, aes(x= t_n)) + 
  scale_color_brewer(palette="Dark2") +  
  geom_vline(data = filter(conc_plot, brk1==1),
             aes(xintercept = t_n, alpha = 0.5), col = "red", linetype = "dashed") +
  geom_vline(data = filter(conc_plot, brk2 ==1), 
             aes(xintercept = t_n, alpha = 0.5), col = "blue") +
  geom_point(data = filter(conc_plot, ServeNo == 1), 
             aes( y = speed, col = as.factor(ServeNo))) + 
  geom_line(data = subset(conc_plot, ss_pred_w_impt >175), 
            aes(y = ss_pred_w_impt,col = as.factor(ServeNo+1))) + 
  theme(legend.position = "none") + 
  ggtitle("First Serve")
 
p2 <- ggplot(conc_plot, aes(x= t_n)) + 
  scale_color_brewer(palette="Dark2") +  
  geom_vline(data = filter(conc_plot, brk1==1),
             aes(xintercept = t_n, alpha = 0.5), col = "red", linetype = "dashed") +
  geom_vline(data = filter(conc_plot, brk2 ==1), 
             aes(xintercept = t_n, alpha = 0.5), col = "blue") +
  geom_point(data = filter(conc_plot, ServeNo == 2), 
             aes( y = speed, col = as.factor(ServeNo))) + 
  geom_line(data = subset(conc_plot, ss_pred_w_impt <175), 
            aes(y = ss_pred_w_impt,col = as.factor(ServeNo+1))) + 
  theme(legend.position = "none") +
  ggtitle("Second Serve")
  
grid.arrange(p1, p2, nrow = 2)
```

```{r}
ss_t = sum((conc_plot$speed - mean(conc_plot$speed))^2)
ss_resid = sum((conc_plot$speed - conc_plot$ss_pred_w_impt)^2)
R_sqr = 1- ss_resid/ss_t
```

R-square = 0.5074

---

## Female data 

- Kerber | 2016 | Ausopen | Final |Against Serena Williams


```{r echo = FALSE, message= FALSE}
Kerber <- Alex <- extract_records_all("data/2016-ausopen-matches.csv", 
                               "data/2016-ausopen-points.csv", "Angelique Kerber")



Kerber <- Kerber %>% filter(match_num == 2701)

k <- data.frame(brk1 = ifelse(rest(Kerber) == 1.5, 1,0), 
                brk2 = ifelse(rest(Kerber) == 2, 1,0), 
                impt = point_impt(Kerber), 
                t = Kerber$PointNumber, 
                match_num = Kerber$match_num,
                speed = Kerber$Speed_KMH,
                GameNo = Kerber$GameNo,
                ServeNo = Kerber$ServeNumber)

k <- k %>% replace_na(list(impt = 0))

ggplot(k) + geom_histogram(aes(x = impt))

k <- k %>% filter(speed != 0 )%>% 
  mutate(t_n = (t-min(t))/(max(t)-min(t))*100) %>% 
  mutate(impt_high = ifelse(impt>=0.1, 1,0)) %>% 
  mutate(firstserve_index = ifelse(ServeNo ==1, 1,0 ))


ggplot(k) + 
  geom_point(aes(x = t_n, y = speed, col = as.factor(ServeNo))) +
  geom_vline(data =filter(k, brk1 ==1), 
             aes(xintercept = t_n, alpha = 0.5), col = "grey", linetype = "dashed") + 
  geom_vline(data = filter(k, brk2 ==1), 
             aes(xintercept = t_n, alpha = 0.5), col = "black", linetype = "dashed") + 
  scale_y_continuous(breaks = c(140,160,180,200, 220)) + 
  theme(legend.position = "none")

```


```{r echo = FALSE}
brk1 <- k$brk1
brk2 <- k$brk2
brk1_index <- which(brk1 %in% 1)
conc_brk1 = data.frame(matrix(nrow = nrow(k), ncol =length(brk1_index)))
brk2_index <- which(brk2 %in% 1)
conc_brk2 = data.frame(matrix(nrow = nrow(k), ncol =length(brk2_index)))


ll <- function(par, data){
  conc_1 = par[4]*(exp(k$t_n*par[1]))
  conc_2 = par[5]*(exp(k$t_n*par[2]))
  conc_3 = (par[5] + par[6])*(exp(k$t_n*par[3]))
  
  for (i in 1: length(brk1_index)){
    conc_brk1[,i] = c(rep(0,brk1_index[i]),conc_2)[1:nrow(k)]}
  conc_brk1_sum <- rowSums(conc_brk1)
  
  for (i in 1: length(brk2_index)){
    conc_brk2[,i] = c(rep(0,brk2_index[i]),conc_3)[1:nrow(k)]}
  conc_brk2_sum <- rowSums(conc_brk2)
  
  ss_pred = par[7] + conc_1 + conc_brk1_sum + conc_brk2_sum+ 
    par[8]*k$impt_high +par[9]*k$firstserve_index
  
  u = k$speed - ss_pred
  f <- -sum(log(dnorm(u, par[10], par[11])))
  return(f)
} 

ll(c(-0.9,-5,-5,10,10,10, 151, 10, 10, 5,20),k)

fit_1_k <- optim(c(-0.05,-10,0,10,5,5, 
                 151, 20, 20, 0,10), 
               ll,method= "L-BFGS-B", data = k,
               lower = c(-0.2, -10, -10, 0,0,0, -Inf, -Inf, -Inf, -Inf,-Inf), 
               upper = c(0,0,0,20,10,10,Inf, 20, Inf, Inf, Inf))


# fit_1_k$par

```


```{r  echo = FALSE, result = FALSE}
conc_1<- fit_1_k$par[4]*exp(k$t_n*fit_1$par[1])
conc_2<- fit_1_k$par[5]*exp(k$t_n*fit_1$par[2]) 
conc_3 <- (fit_1_k$par[5] + fit_1_k$par[6])*exp(k$t_n*fit_1_k$par[3]) 

conc_exp_k <- cbind(k, conc_1, conc_2,conc_3)

p1 <- ggplot(conc_exp_k) + geom_line(aes(x = t_n, y = conc_1)) 
p2 <- ggplot(conc_exp_k) + geom_line(aes(x = t_n, y = conc_2)) + 
  scale_x_continuous(limit = c(0, 10), breaks = c(0,2,4,6,8,10))
p3 <- ggplot(conc_exp_k) + geom_line(aes(x = t_n, y = conc_3)) +
  scale_x_continuous(limit = c(0, 10), breaks = c(0,2,4,6,8,10))
gridExtra::grid.arrange(p1,p2,p3, nrow = 3)




# compute lagged expoential for breaks
for (i in 1: length(brk1_index)){
  conc_brk1[,i] = c(rep(0, brk1_index[i]),conc_2)[1:length(conc_2)]
}
conc_brk1_sum <- rowSums(conc_brk1)


for (i in 1: length(brk2_index)){
  conc_brk2[,i] = c(rep(0, brk2_index[i]),conc_3)[1:length(conc_3)]
}
conc_brk2_sum <- rowSums(conc_brk2)


conc_plot_k<- cbind(conc_exp_k, conc_brk1_sum, conc_brk2_sum)
conc_plot_k <- conc_plot_k %>% 
  mutate(base = fit_1_k$par[7] + conc_1 +fit_1_k$par[9]*k$firstserve_index) %>% 
  mutate(base_brk1 = fit_1_k$par[7] + conc_1 +conc_brk1_sum + 
           fit_1_k$par[9]*k$firstserve_index) %>% 
  mutate(base_brk2 = fit_1_k$par[7] + conc_1 +conc_brk2_sum + 
           fit_1_k$par[9]*k$firstserve_index) %>% 
  mutate(ss_pred_no_impt = fit_1_k$par[7] + conc_1 +conc_brk1_sum + conc_brk2_sum + 
           fit_1_k$par[9]*k$firstserve_index) %>% 
  mutate(ss_pred_w_impt = fit_1_k$par[7] + conc_1 +conc_brk1_sum + conc_brk2_sum + 
           fit_1_k$par[8]*k$impt_high + 
           fit_1_k$par[9]*k$firstserve_index)  

```


---

# Result


```{r echo = FALSE, message=FALSE, fig.height=7, fig.width=10, dev="svg"}
p1 <- ggplot(conc_plot, aes(x= t_n)) + 
  geom_line(data = subset(conc_plot, ss_pred_w_impt >175), aes(y = base), linetype = "dashed")+
  geom_line(data = subset(conc_plot, ss_pred_w_impt <175), aes(y = base), linetype = "dashed")+
  geom_line(data = subset(conc_plot, ss_pred_w_impt >175), aes(y = ss_pred_w_impt)) +
  geom_line(data = subset(conc_plot, ss_pred_w_impt <175), aes(y = ss_pred_w_impt)) +
  geom_vline(data = filter(conc_plot, brk1==1),
             aes(xintercept = t_n, alpha = 0.5), col = "red", linetype = "dashed") +
  geom_vline(data = filter(conc_plot, brk2 ==1), 
             aes(xintercept = t_n, alpha = 0.5), col = "blue") + 
  theme(legend.position = "none") + 
  xlab("t") + ylab("Speed_pred") + 
  ggtitle("Alex")

p2 <- ggplot(conc_plot_k, aes(x= t_n)) + 
   geom_line(data = subset(conc_plot_k, ServeNo ==1), aes(y = base), linetype = "dashed")+
  geom_line(data = subset(conc_plot_k, ServeNo ==2), aes(y = base), linetype = "dashed")+
  geom_line(data = subset(conc_plot_k, ServeNo ==1), aes(y = ss_pred_w_impt)) +
  geom_line(data = subset(conc_plot_k, ServeNo ==2), aes(y = ss_pred_w_impt)) +
  geom_vline(data = filter(conc_plot_k, brk1==1),
             aes(xintercept = t_n, alpha = 0.5), col = "red", linetype = "dashed") +
  geom_vline(data = filter(conc_plot_k, brk2 ==1), 
             aes(xintercept = t_n, alpha = 0.5), col = "blue") + 
  theme(legend.position = "none") +
  xlab("t") + ylab("Speed_pred") + 
  ggtitle("Kerber")

gridExtra::grid.arrange(p1, p2, nrow = 2)
```



---

class: center, middle

# Thanks!

Supervised by Di Cook and Stephanie Kovalchik



