---
title: "Dose Response Model"
author: "Huize Zhang"
date: "23/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(drc)
library(sandwich)
library(lmtest)
library(tidyverse)

Federer = extract_records_all("../data/2017-ausopen-matches.csv","../data/2017-ausopen-points.csv", "Roger Federer")
```

```{r}
# reproduce Steph's work
# fatigue_predict <- function(data, serve, the_game){
#   # data <- subset(data, ServeNumber == serve & game == the_game)
#   
#   fit <- drm(log(Speed_KMH) ~ Set, 
#              data = data,
#              fct = W1.4())
#   
#   data.frame(
#     game = paste("game", data$game[1]),
#     ServeNumber = ifelse(serve == 1, "First Serve", "Second Serve"),
#     Speed = c(data$Speed_KMH, exp(predict(fit))),
#     Type = rep(c("Actual","Predicted"), each = nrow(data))
#   )
# }	
# 
# params <- expand.grid(game = unique(f$game), serve = 1:2)
# model <- fatigue_predict(f,params$serve,params$game)

```

The current problem faced when fitting a dose-response model is that the evidence of fatigue (reduction of serving speed) is not very obvious. For the three set of data manipulated below, only the second df produce a downward sloping dose-response curve. 

The last df, which is group_by SetNo fitts the dose-response model quite nice (dr4pl)
```{r}
df <- Federer %>% group_by(game) %>% 
  dplyr::select(game,Speed_KMH) %>% 
  summarise(mean_speed = mean(Speed_KMH)) 

# df<- Federer %>% group_by(game) %>% 
#   filter(SetNo ==5) %>% 
#   dplyr::select(game,Speed_KMH) %>% 
#   filter(Speed_KMH != 0)%>% 
#   summarise(mean_speed = mean(Speed_KMH)) %>% 
#   filter(game > 150 & game <240) %>% mutate(game = game - 213) 
# 
# df <- Federer %>% group_by(game) %>% 
#   filter(SetNo ==5) %>% 
#   dplyr::select(game,Speed_KMH) %>% 
#   filter(Speed_KMH != 0)%>% 
#   summarise(mean_speed = mean(Speed_KMH)) %>% 
#   filter(game > 250) %>% mutate(game = game - 257)

df <- Federer %>% group_by(SetNo) %>% 
  dplyr::select(SetNo, Speed_KMH) %>% 
  filter(Speed_KMH != 0) %>% 
  summarise(mean_speed = mean(Speed_KMH)) %>% 
  mutate(SetNo = as.numeric(SetNo)) %>% 
  mutate(mean_speed = as.numeric(mean_speed))
df
```

Dose-response model via drc package. If meet convergence problem, use dr4pl
```{r}
model <- drm(df$mean_speed~df$SetNo, fct = LL.4())
summary(model)
# adjust the std. error by sandwich package
coeftest(model, vcov = sandwich)
ED(model, 50, interval = "delta")
plot(model, broken = TRUE, type = "all", xlim = c(1, 20))
plot(df$SetNo, predict(model))

```

dr4pl also produces dose-response model as drc package. It sometimes works better than drc when there's convergence problem with drc
```{r}
library(dr4pl)
m <- dr4pl(response = df$mean_speed, dose = df$SetNo, method.init = "logistic", trend = "decreasing")
summary(m)
plot(m)
values <- IC(m, c(95,90,80,50))
values

```

Dose Response model only allows for one x variable dose (although it allows for multiple response value for one particular dose). 
We want to incorporate in more variables (rest, running distance) to model the spped_KMH, so move to the glm model. 
-- still building ---
```{r}
f_dose <- dose(Federer %>% 
                 filter(match_num != 1601) %>% 
                 filter(Speed_KMH !=0))

df <- Federer %>% filter(match_num != 1601) %>% 
  filter(Speed_KMH != 0) %>% 
  mutate(rest = cumsum(f_dose)) %>% 
  mutate(dist= ifelse(player1 ==" Roger Federer", P1DistanceRun, P2DistanceRun)) %>% 
  mutate(distance = cumsum(dist)) %>% 
  dplyr::select(round, SetNo, game, Speed_KMH,rest, dist, distance) 
df

# could use sum(distance)
game <- df %>% group_by(game) %>% 
  summarise(speed = mean(Speed_KMH), rest = max(rest), distance = max(distance))

set <- df %>% group_by(SetNo) %>% 
  summarise(speed = mean(Speed_KMH), rest = max(rest), distance = max(distance))

match <- df %>% group_by(round) %>% 
  summarise(speed = mean(Speed_KMH), rest = max(rest), distance = mean(distance))


library(visreg)
fit <- glm(speed ~rest + distance,data = match,family = Gamma())
summary(fit)
visreg(fit, "distance", gg=TRUE) + theme_bw()
anova(fit,"chisq")

```


